Lightweight Scheduling for the PRAGMA Cloud Testbed

### 2016-07-15 paper.tex
More changes to address reviewer comments -- also added Nadya's change

Booked~\cite{booked} from Twinkle Toes software seemed the best suited to our needs compared to all the calendar reservation systems we evaluated.  It was easy to setup, had  a good look and feel, and was customizable.  It also had some nice additional features like usage reporting, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions management, and user and group quotas.  After working with Booked during our pilot setup, we did find some weaknesses.  First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure.  For example, we had to go through many source code files to find the right place to change  or add functionality.   Second, the reservation frequency of the event is currently limited to one hour.  Based on today's PRAGMA needs this is sufficient because the time it takes to start up the virtual cluster 
is in the order of tens of minutes and  the user reservation is expected to be in the order of days. In the future when the virtual cluster startup  time shortens to minutes, the reservation frequency granularity can be optimized in order to avoid a potential of the underutilized resources.

@@

\begin{enumerate}
\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (e.g., IPOP, ViNE~\cite{vine}, or PRAGMA-ENT) that is needed.  
\item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  
\item The user selects the desired resources, the timeframe, virtual cluster image, and configuration details and submits them to the Cloud scheduler.  Once the reservation is verified and confirmed, the user will get an email confirmation.  In the event two users try to reserve the same resources at the same time,
\item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when the reconfigured cluster is ready.   
\end{enumerate}

@@

\begin{itemize}
\item Added the ability to make multiple reservations on a resource per time slot to allow more than one virtual cluster to run simultaneously on a site.   This was the most extensive change we made to Booked.  Fortunately, no changes were needed to the backend database but it required several changes to the GUI code to allow an additional reservation to be made on a resource and to display multiple reservations per time slot in the calendar views.  In total, enhancements were made to seven PHP files and two template files.
\item Added a custom field for a public SSH key  to the user profile.  When PCC launches a virtual cluster, it pulls the public SSH key from the user's  profile and gives it to pragma\_boot so that the user can login as root once the virtual cluster is launched.   The ability to use a root login is not required for all but can be needed in some case if users want to do system modifications. A user with a root login has a full ownership of the virtual cluster. 
\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to the reservation specification.
\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT capability (enabled, disabled) to the resource specification.
\item Added a numeric count as a custom field type.  Booked only provided text matching on search so if a host had 16 CPUs available and the user requested 8 CPUs, the search would fail per "16" != "8" inequality.  By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed for "8 \textless= 16".
\item Added virtual cluster statuses: "Starting", "Running", and "Stopping".  Each virtual cluster status is shown as a different color in  the calendar view.  "Starting" means it has been approved and is being started up on the selected hypervisor... SHAVA EXPAND THIS
\item Added the ability to retrieve and set the reservation status from the Booked REST API.  	
\item Added the PRAGMA logo to the header.
\end{itemize}

@@

@@

In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.  NADYA TO PUT SENTENCES FOR REVIEWER 1.  In the future when we deploy on more sites, we can gather schedule statistics on usability and applicability of different scenarios.   We will also work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.  Finally, we will investigate using the CloudFront option in pragma\_boot to manage application virtual cluster images and staging them to each of the sites.     We will package and document the resulting software so that new users and sites can get quickly onto the PRAGMA cloud testbed and use it as an experimental service to test new science and infrastructure tools.

>>>>

Booked~\cite{booked} from Twinkle Toes software seemed the best suited to our needs compared to all the calendar reservation systems we evaluated.  It was easy to setup, had  a good look and feel, and was customizable.  It also had some nice additional features like usage reporting, support for different time zones for each user, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions management, and user and group quotas.  After working with Booked during our pilot setup, we did find some weaknesses.  First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure.  For example, we had to go through many source code files to find the right place to change  or add functionality.   Second, the reservation frequency of the event is currently limited to one hour.  Based on today's PRAGMA needs this is sufficient because the time it takes to start up the virtual cluster 
is in the order of tens of minutes and  the user reservation is expected to be in the order of days. In the future when the virtual cluster startup  time shortens to minutes, the reservation frequency granularity can be optimized in order to avoid a potential of the underutilized resources.

@@

\begin{enumerate}
\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (e.g., IPOP, ViNE~\cite{vine}, or PRAGMA-ENT) that is needed.  
\item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  
\item The user selects the desired resources, the timeframe, virtual cluster image, and configuration details and submits them to the Cloud scheduler.  Once the reservation is verified and confirmed, the user will get an email confirmation.  Otherwise if there is a problem with the reservation such as someone else has taken the resources while the user was filling in the reservation fields, the user will get notified immediately of the failure once they hit the submit button or later via email.
\item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when the reconfigured cluster is ready.   
\end{enumerate}

@@

\begin{itemize}
\item Added the ability to make multiple reservations on a resource per time slot to allow more than one virtual cluster to run simultaneously on a site.   This was the most extensive change we made to Booked.  Fortunately, no changes were needed to the backend database but it required several changes to the GUI code to allow an additional reservation to be made on a resource and to display multiple reservations per time slot in the calendar views.  In total, enhancements were made to seven PHP files and two template files.
\item Added a custom field for a public SSH key  to the user profile.  When PCC launches a virtual cluster, it pulls the public SSH key from the user's  profile and gives it to pragma\_boot so that the user can login as root once the virtual cluster is launched.   The ability to use a root login is not required for all but can be needed in some case if users want to do system modifications. A user with a root login has a full ownership of the virtual cluster. 
\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to the reservation specification.
\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT capability (enabled, disabled) to the resource specification.
\item Added a numeric count as a custom field type.  Booked only provided text matching on search so if a host had 16 CPUs available and the user requested 8 CPUs, the search would fail per "16" != "8" inequality.  By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed for "8 \textless= 16".
\item Added virtual cluster statuses: "Starting", "Running", and "Stopping".  Each virtual cluster status is shown as a different color in  the calendar view.  "Starting" means it has been accepted by the scheduler and is being started up on the local cloud systems, "Running" means that the resources are available for the user to log into, and "Stopping" means the virtual cluster is being shutdown and is no longer available for log in.
\item Added the ability to retrieve and set the reservation status from the Booked REST API.  	
\item Added the PRAGMA logo to the header.
\end{itemize}

@@

We still need to address the scenario where two users might try to reserve the same resources at the same time as a race condition is possible.  Booked does not have are not any locking mechanisms in place currently that would handle two simultaneous user requests for the same resources gracefully and so this will need to be added in the future.    We also are not handling hardware failures at resource sites so when a  virtual cluster goes down; the failure would be recorded but we cannot accommodate relocation.  For fault tolerance of the cloud scheduler, we will deploy a backup instance somewhere else so that we can failover in the event of an outage.

@@

In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.  Currently,  the startup latency from a virtual cluster reservation to its availability to the user has not been analyzed in detail but we observed it is on the order of tens of minutes for our pilot implementation.   This user latency has multiple factors such as the underlying virtualization manager and load on the physical system where the virtual cluster is instantiated. There is additional latency if the cluster image is not available on the site as it will need to be downloaded from a remote repository and processed to make it available on the site. While this applies only for the first virtual cluster instantiation, it can add a considerable time  latency  depending on the cluster image size and the network speed from the site to the remote repository.

In the future when we deploy on more sites, we can gather statistics on the cloud scheduler's usability and user experiences with virtual clusters startup latencies in addition to its applicability to different scenarios.   We will also work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.  Finally, we will investigate using the CloudFront option in pragma\_boot to manage application virtual cluster images and staging them to each of the sites.     We will package and document the resulting software so that new users and sites can get quickly onto the PRAGMA cloud testbed and use it as an experimental service to test new science and infrastructure tools.
